Tutkielman otsikko

### 2020-02-02 gradu/msc_english_contents.tex
Adds more text

\chapter{Basic Concepts}
Variable-byte encoding \citep{Wil99} is a method of compressing integers. It splits an integer into blocks of $b$ bits and then encodes
the block as $b + c$. $c$ is a bit denoting whether $b$ has the most significant bits of the number or not. A good dataset for variable-byte encoding
is a set with both very small and very large numbers, since he compression comes from omitting the leading zeros in the bit array. This method 
loses in compression performance to other methods \citep{Bri09}, but they are very fast to decode and allow fast random access.

- Explain rank / select

\chapter{Previous Work}
bl
\chapter{Algorithm}
here
\chapter{Results}
hmll
\section{Figures}
Figure~\ref{fig:logo} gives an example how to add figures to the document. Remember always to cite the figure in the main text.

>>>>

\chapter{Basic Concepts}
Variable-byte encoding \citep{Wil99} is a method of compressing integers. It splits an integer into blocks of $b$ bits and then encodes
the block as $b + c$. $c$ is a bit denoting whether $b$ has the most significant bits of the number or not. (example?) A good dataset for variable-byte encoding
is a set with both very small and very large numbers, since the compression comes from omitting the leading zeros in the bit array. This method 
loses in compression performance to other methods \citep{Bri09}, but decoding is fast.

Rank and select are two functions that work on bit arrays. Rank(i) gives the sum of 1 bits from the beginning of the bit array and select(i) gives
the index of ith 1 bit in the bit array. Both functions work in constant time (citation?) and they require only a few percents of extra space over
the data. The extra bits $c$ added by variable-byte encoding conveniently create a bit array, where 1's represent the endings of numbers. An effective
version of random access has already been introduced \citep{Bri09}. It uses one rank call for each block after the first which makes the fetching very
fast if the number in question fits to one block.

An effective version with select is also possible. The index of the block where ith number starts is obtainable with select(i-1)+1, when i > 0 and 0 when i=0.
By separating the $c$ bit array and $b$ block array, $b$ now contains variable-byte integers in readable form. Another upside is that functions 
next(i) and previous(i) are conveniently available without any extra calculation and the runtime is not dependent on the size of the value.

- explain how random access is good

\chapter{Previous Work}
bl

\chapter{Algorithm}
 - modifications to basic implementation

\chapter{Results}
 - comparison to basic implementation + Bri09
 - 
\chapter{Conclusion}
 - here

\chapter{Future work}
 - something to improve / research?

\section{Figures}
Figure~\ref{fig:logo} gives an example how to add figures to the document. Remember always to cite the figure in the main text.
